\vspace*{0.1cm}
\begin{center}
    {\LARGE \textsc{Acknowledgments}}
\end{center}
\vspace{0.3cm}

%\begin{flushright}
%\begin{tabular}{l}
%"\textit{The time has come,}" \textit{the Walrus said,}\\
%"\textit{To talk of many things}" \dots \\
%\hspace{1in} - {\scriptsize Lewis Carroll, Through the Looking-Glass}\\
%\end{tabular}
%\end{flushright}

We would like to extend our heartfelt appreciation to our thesis supervisor, 
\textbf{Dr. SK Hafizul Islam}, Assistant Professor, Department of Computer Science and Engineering, 
Indian Institute of Information Technology Kalyani, for his invaluable mentorship, consistent encouragement, 
and unwavering support throughout the duration of our research. His thoughtful insights and constructive feedback 
have been instrumental in shaping the direction and quality of this work.
Our sincere thanks go to the faculty and staff of the Department of Computer Science and Engineering and the 
Department of Electronics and Communications Engineering for fostering an academically enriching environment 
and providing essential facilities and resources.

We owe our deepest gratitude to our families and friends for their patience, motivation, 
and emotional support, which helped us remain focused and determined throughout this journey.

Lastly, we would like to thank everyone who contributed, directly or indirectly, to the successful completion 
of this thesis.

\vspace{1in}

\begin{minipage}[l]{0.2\textwidth}%
    \textbf{Date:}~
    \\
    \\
    \\
    \\
    \\
    \vspace{2cm}
\end{minipage}
\hfill
\begin{minipage}[r]{0.7\textwidth}%
    \begin{flushright}
         \textbf{Subhradeep Ray (Reg. No.: CSE/21090/750)}
        \text{Department of Computer Science and Engineering}\\
        \text{Indian Institute of Information Technology Kalyani}\\
        \text{West Bengal - 741235, India}\\
        \vspace{3cm}
        \textbf{Viswanath Odiya (Reg. No.: ECE/21157/817)}\\
        \text{Department of Electronics and Communication Engineering}\\
        \text{Indian Institute of Information Technology Kalyani}\\
        \text{West Bengal - 741235, India}
    \end{flushright}
\end{minipage}
